Community biomass and species richness temporal trends reveal contrasting assembly and disassembly processes in natural food-webs

### 2022-11-08 paper/manuscript.Rmd
Finalise manuscript preparation

---
title: "Community biomass and species richness temporal trends reveal contrasting assembly and disassembly processes in natural food-webs"
author: Alain Danet, Loubna El Madouri, Maud Mouchet, Elisa Thébault and Colin Fontaine
date: \today
output:
#  bookdown::word_document2:
#    fontsize: 12pt
  bookdown::pdf_document2:
    fig_caption: true 
    keep_tex: true
    toc: false 
    fontsize: 12pt
header-includes:
   - \usepackage{natbib}
   - \usepackage{float}
   - \usepackage{multirow}
   - \usepackage{booktabs}
   - \usepackage{makecell}
   - \usepackage{setspace}
   - \doublespacing
   - \usepackage{lineno}
   - \linenumbers
geometry: margin=2cm
number_sections: true
bibliography: references.bib
---

@@

# Abstract (267 / 250 words max)

Understanding how biodiversity changes is a central question in ecology, but
still challenging because of the multidimensionality of biodiversity changes.
Although the structure of ecological networks is a crucial determinant of
community dynamics, this dimension of biodiversity changes have been greatly
overlooked. One main reason is the scarce availability of the temporal series of
ecological networks. Our knowledge of the relationships among structural
properties of ecological networks comes mainly from spatial comparison, assuming
a space-for-time equivalence. However, little is know about the validity of the
space-for-time hypothesis, i.e. if spatial comparison of ecological networks
accurately translates temporal sequences of assemblages. Here we questioned the
relationships between the temporal trends of food-web structure and the ones
of community biomass and species richness, and we assessed how those
relationships differed from the ones estimated in space. We inferred the
structure of using food diet and individual body size data from fish monitoring
in France. We estimated the temporal trends of food-web structure in 400 sites
using a hierarchical Bayesian model and then assess the relationships among them
using structural equation models. We found that temporal trends of connectance
and average trophic level are positively related to ones of community biomass
but surprisingly poorly related to the temporal trends of species richness. We
further found support for space-for-time hypothesis for community biomass, but
not for species richness. We suggest that our results resonates with the
energetic paradigm of food-webs. Our study sheds a new light on the dynamics of
food-web assemblage in natural settings and how the structure of ecological
networks changes according to the changes of other biodiversity facets.

# Significance (109 / 120 words max)

While the multidimensional nature of biodiversity changes has been clearly
emphasized, little is know how the structure of food-web changes according to
changes in species richness and community biomass. By inferring food-web
structure based on food-diet data and body size ratios, we analyzed how food-web
structure in time according to simultaneously changes in species richness and
community biomass. We identify that community biomass is the primary driver of
temporal changes in foodweb structure, but that species richness has surprisingly
low effect. We further show that space-for-time leads to contrasting results for
species richness. Our results shed a new light on how foodweb assemble
and disassemble in natural settings.

@@

There is rising recognition that biodiversity changes are
multidimensional [@mcgill_fifteen_2015; @blowes_geography_2019;
@hillebrand_biodiversity_2018; @blowes_local_nodate]. Studies found important
changes in species composition despite no changes in species richness in average
[@dornelas_assemblage_2014; @blowes_geography_2019], and highlighting the need
for further investigations on how biodiversity changes over time.

Yet species interactions are largely absent from biodiversity changes. Although
species interactions, described as ecological networks, drive for a large part
community composition, species coexistence, and ecosystem
functioning [@chesson_mechanisms_2000; @thebault_food-web_2003;
@oconnor_general_2017]. The structure of ecological networks can trigger or
impede perturbation propagations, i.e secondary extinctions [@paine_note_1969;
@dunne_network_2002; @srinivasan_response_2007;
@stouffer_compartmentalization_2011]. Higher connectance, which measures the
degree of generalism in a network, is predicted to increase robustness to
secondary extinctions [@dunne_network_2002; @sole_complexity_2001] and studies
found that higher trophic level was associated to a lower temporal variability
of community biomass in theoretical models and empirical
data [@shanafelt_stability_2018; @danet_species_2021]. Then, accounting for how
the structure of ecological networks changes over time according to other
biodiversity facets is crucial.

Th availability of ecological network timeseries is
scarce [@kortsch_disentangling_2021], such as predictions about how the structure
of ecological networks change according to changes in species richness come from
spatial comparison of networks. Such studies showed that connectance is
negatively related to species richness in small (i.e. below 20 species)
empirical food-webs [reviewed in @dunne_network_2006], which reflects
constraints on the number of realizable species interactions or an increasing
proportion of specialists in more rich foodwebs. Connectance has been shown to
scale positively or negatively with community biomass in theoretical
food-webs [@thebault_food-web_2003; @poisot_trophic_2013] while the average
trophic level was previously shown to scale negatively with species
richness [@danet_species_2021].

However, predicting how ecological networks will change in time according to
other biodiversity facets based on spatial comparison of ecological communities
is valid as long as assumption of the space-for-time substitution is valid as
well. Space-for-time substitution supposes a homology between variations in
space and in time, i.e. that spatial sequences correspond to temporal sequences.
This method have raised criticisms [@damgaard_critique_2019], with studies
showing analogous patterns in space and time [@blois_space_2013] or not
[@johnson_testing_2008; @adler_contrasting_2007]. Spatial turnover of species
composition, dispersal limitation, disturbance and stochastic processes have
been hypothesised to be responsible for the non-compliance of space-for-time
substitution [@adler_contrasting_2007; @blois_space_2013].

@@

Using a hierarchical model to estimate conjointly the temporal trends of  
foodweb structure at all sites, we found weak evidence of a global decrease in
community biomass (CI80%:`r gbm` by decade; median [Credible Interval]) but of a
global increase in species richness (CI80%: `r gr` by decade) across the `r nb_st` sites. 
We did not find evidence of a global trends in connectance and average trophic
level (CI80%: `r gct` and `r gt` respectively). The hierarchical model captured
well the variance associated to the spatial structure of the data with
conditional $R^2$ going from  `r r2ic$log_bm_std` (mean [CI95%]) for community
biomass to `r r2ic$log_rich_std` for species richness (Table S1).

@@

Using structural equation models to disentangle the direct and indirect
relationships among foodweb structure descriptors, we found a strong positive
relationship between the temporal trends of community biomass and of
species richness ($r_\delta =$ `r rdtps$bm$r`, Fig. \@ref(fig:sem) a and b,
$r_\delta$: standardized slope coefficients, [95% CI]), meaning that both
temporal trends had the same direction in average. At the sites where species
richness decreased over time, community biomass also decreased over time. On the
other hand, at the sites where species richness increased over time, community
biomass also increased over time.

@@

We found that both connectance and average trophic level had a positive
relationship with community biomass over time, meaning that connectance and
average trophic level increase over time when community biomass increases over
time as well, and likewise that connectance and average trophic level decrease
over time when community biomass decreases over time. We also found that the
temporal trends of connectance and average trophic level were strongly
correlated (Fig. S2). All together, it suggests that the disturbances that
lead to decline or increase of community biomass over time are concomitant to
the decline and increase respectively of trophic species at high trophic level.
Likewise, we found that the temporal trends of community biomass were positively
associated to the temporal trends of both number and proportion of piscivorous
fish species present in the communities (Fig. S2). This result is in
accordance with the energetic paradigm of food-webs which proposes that
food-webs are primarily constrained by fluxes of matter and energy
[@lindeman_trophic-dynamic_1942], which suggest that productivity constraints
food chain length by energy limitation [@jenkins_productivity_1992;
@takimoto_ecosystem_2008]. Higher trophic levels have been shown to be more
sensible to ecosystem size [@takimoto_ecosystem_2008; @post_ecosystem_2000], to
resource availability and disturbance in small systems [@post_long_2002] and
particularly to disturbance in hydrologic regimes in streams [@sabo_role_2010;
@mchugh_dual_2010; @jake_vander_zanden_global_2007]. In the same line, the
metabolic theory of ecology also predicts that top predators are highly
constrained by energy limitation [@arim_relationship_2007]. However this
interpretation does hold if the temporal trends of community biomass are related
to the quantity of resource available. Another explanation can be that top
trophic level species are often the ones that are targeted, harvested and those
that are artificially introduced for recreational fisheries
[@reid_emerging_2019]. Further investigations on the drivers of community
biomass decline and increase over time would be needed to further conclude.

In contrast, we found that the temporal trends of connectance was negatively
related to the temporal trends of species richness, in accordance with the
strong negative relationships that we found in the present study and from
previous results over spatial gradient [@dunne_network_2006;
@danet_species_2021]. Furthermore, we found that temporal trends of the
proportion of piscivorous species was negatively associated with the temporal
trends of species richness (Fig. S2), meaning that communities that increase in
richness over time have an increase in the proportion of non-piscivore over time
and vice versa. This result is in accordance with a meta-analysis showing the
number of species by trophic level decrease toward higher trophic level
[@turney_pyramids_2016]. It suggests that the temporal increase of species
richness found in our dataset is associated with a decrease of connectance
because species richness increases are biased toward non-piscivorous species
which have no piscivory links. We also found that the changes of average trophic
level over time were not related to the changes of species richness over time.
All together, an explanation can be that species richness changes over time,
without changes in community biomass, are more linked to non piscivorous species
having lower trophic level and relatively low biomass, which results in negative
relationships with temporal trends in connectance and no relationship with
average trophic level.

However, as we found a strong positive relationship between the temporal trends
of species richness and community biomass, it results that species richness
temporal trends had an indirect positive effect on the temporal trends of both
connectance and average trophic level via community biomass. Then, our study
suggests that species richness and community biomass have antagonistic effects
on connectance, respectively negative and positive. The temporal trends of
species richness had then a total effect on the temporal trends of connectance
and average trophic level which was twice lower than the one of the temporal
trends of community biomass. It suggests that the temporal assembly and
disassembly of food-webs are more sensitive to variations in community biomass
than in species richness, in other words that food-web structure is more robust
to changes in species richness than in biomass over time. Our study then
suggests that the changes of food-web structure over time related to changes in
species richness could be actually mediated by changes in community biomass.

As studies reported community biomass or species richness temporal trends
[@hallmann_more_2017; @ceballos_biological_2017; @dornelas_assemblage_2014;
@blowes_geography_2019], we document that temporal trends of species richness
and community biomass might have different consequences on the structure of the
community. In particular, the decrease of the average trophic level associated
to the loss of community biomass could have cascading effect on ecosystem
functioning as the highest predators are predicted to increase productivity and
stabilize communities [@wang_biodiversity_2018; @shanafelt_stability_2018;
@danet_species_2021]. In contrast, the loss of species richness might be linked
to loss of redundancy in the food-web, highlighting that those communities may
become more at risk regarding future perturbations [@yachi_biodiversity_1999].
Interestingly, the hypothesis that temporal trends in community biomass is
related to higher trophic level in our system has a strong reasoning with the
effect of human fisheries in marine systems which affect primarily fish species
at top trophic levels [@sprules_surfing_2016].

Regarding the space-for-time substitution, we found that connectance was
positively linked to community biomass both over space and time. This result
resonates with theoretical work showing that higher food-web connectance can
result in higher community biomass [@thebault_food-web_2003], or lower
[@poisot_trophic_2013] over space, highlighting that the relation between
connectance and community biomass may be dependant on how food-web structure
change with community biomass over space and time. Likewise, we found that the
average trophic level had a positive relationship with community biomass over
space and time, in accordance with a theoretical study over space predicting an
increase of community biomass subsequent to the presence of higher predators
[@wang_biodiversity_2018]. Finally, we found a positive relationship between
community biomass and species richness both in space and time, however the
temporal relationship was stronger. We also found a negative relationship
between connectance and species richness space and time, in accordance with
previous findings over space in empirical food-webs [@dunne_network_2006]. We further
found no relationship between average trophic level and species richness over
time while we found a twice stronger but positive relationship between median
average trophic level and median species richness over the spatial gradient. An
explanation might be that environmental gradient and the pattern of fish
migration since the last glaciation, for the strong negative effect of species
richness on both connectance and average trophic level over spatial gradient.
However, we controlled our spatial analysis by the Biological Demand in Oxygen,
water temperature and the stream gradient. This result is also in line with
other studies which found a mismatch between spatial and temporal patterns
[@adler_contrasting_2007; @johnson_testing_2008]. This discrepancy in space
and time might be due to the strong assumption that variations in space reflect
differences in assembly stages of ecological communities and then mirror their
temporal assembly/disassembly trajectories. Temporal variation of species
richness might relate to transitory dynamics while spatial gradients reflect
aggregated dynamics over a longer time scale including a far number of processes
including high species turnover along spatial gradients.

@@

Our study suggests that species richness and community biomass temporal trends
in natural fish communities relate on different ecological processes, biomass
temporal dynamics being related to variations of food-web structure at the
higher trophic levels while variations of species richness is weakly related to
change in food-web structure. Our result also suggests that space for time
substitution may be relevant to predict the effects of community biomass losses
or gains on food-web structure but also that space-for-time substitution might
not work for species richness. Our study calls for further investigations to
support the generality of assembly rules in food-webs. They are crucial to
better understand the assembly and disassembly of food-webs and then predict its
impact on ecosystem functioning.

@@

We selected the sites that have been monitored at least 10 years on the
1995-2018 period, with the same sampling protocol over time. We selected
one sampling a year by site. In each site, we ensured low temporal heterogeneity
in sampling period by selecting sampling that were performed in the same period,
leading to a mean standard deviation of sampled month in each site of three
weeks [@danet_species_2021]. Those steps leads to the selection of 
`r nb_st` sites (Figure \@ref(fig:matmet)A).

@@

We first inferred the body mass of each fish individual considering an allometric
law, $B_i = 0.01 \times l_i^{3.03}$, $B_i$ being the biomass expressed in gram and $l_i$ the
body length of the individual $i$ expressed in millimetres [@froese_cube_2006]. Then the
biomass of each species and community biomass were obtained by summing the body
mass of individuals. Finally, the biomass was reported to the sampled surface
(in $m^2$) to account for sampling effort.

@@

We determined the feeding interactions of trophic species according to their
species identity and their body size. The diet database described two or three
life stages by fish species, each one being delimited by body size
[@bonnaffe_comparison_2021; @danet_species_2021]. The ontogenic fish diet
database was filled according to fishbase (Froese et al. 2021) and literature
[@bonnaffe_comparison_2021; @danet_species_2021].

@@

At each site, yearly food-webs were described by their community biomass,
species richness and the evenness of biomass across species. The evenness was
quantified with the Pielou index, computed as the Shannon diversity index
divided by the number of species ($P = \frac{H'}{N}$, with $H'$ being the
Shannon diversity index and $N$ the number of species). Yearly food-web
structure was described by their connectance, describing the degree of
generalism in a network [@dunne_network_2006; @delmas_analysing_2019]. It was
computed as the ratio between the number of actual links and the number of
potential links ($C = \frac{L}{S(S-1)}$, with $L$ being the number of links and
$S$ the number of trophic species). We also computed the average trophic level,
weighted by the biomass of the trophic species. The trophic level of a trophic
species was computed recursively as 1 added to the mean trophic level of the
preys that it consumes [@herendeen_ecological_2008].

@@

The model parameters were estimated using INLA (Integrated Nested Laplacian
Approximation) bayesian inference with default uninformative priors
[@rue_bayesian_2017]. The prior distributions for the fixed intercept and slope were
modelled as a flat normal distribution with a mean of zero ($\mathcal{N}(0,\, 1000)$).
Prior distributions for the gaussian error $\epsilon_{i|n, t}$ and the random
effects were modelled as a log gamma distribution with
shape and inverse scale parameters ($\mathcal{G}(s, \, \tau) = \mathcal{G}(1, \,
5.10^-5)$), then their respective standard deviation were obtained by
backtransforming the inverse scale parameter ($\sigma = 1 / \sqrt{\tau}$).
Community biomass and species richness were logged transformed.

The performance of the model was assessed visually by plotting the fitted versus
the observed values (Fig. SXX). We also visually inspected the Posterior
Integral Transform (PIT) and Conditional Predictive Ordinate (CPO) to assess the
quality of the model fit. We also computed the marginal ($R^2_m$)and conditional
($R^2_c$) R-squared of the models following previous formulations
[@nakagawa_general_2013; @lahuis_explained_2014]. In order to account for the
variability associated with the priors [@gelman_r-squared_2019], we computed the
variance of the posterior distribution  of each predicted values ($Var_{fit} =
\sigma^2(\hat{Y_i})$). We computed $R^2$ values for each predicted values, then
obtaining a $R^2$ distribution [@gelman_r-squared_2019]. We reported mean
marginal and conditional $R^2$ associated the 95% credible interval computed
using the Highest Posterior Density method.

@@

The temporal trends at each site have been extracted from the hierarchical model
using mean BLUP values (Best Linear Unbiaised Prediction), then multiplied
by 10 to get temporal trends by decade. For logged transformed biomass and
species richness, we derived percentage change by decade by transforming the
slopes ($\beta\prime = (e^{\beta } - 1) \times 100 \times 10$).

@@

To assess the validity of the space-for-time substitution hypothesis, we
compared the relationships among food-web structure in time to the ones observed
in space. To do so, we selected the same sites that were used
for the temporal analysis. We summarised temporal series of each station by the
median value over time of each facet of food-web structure described above.

@@

We modelled the relationships among food-web structure, that is
connectance and average trophic level, and community biomass ($dB$) and species
richness ($dR$) with Structural Equation Models. We modelled a it such as species
richness is related to community biomass and food-web structure. In
turn community biomass was related to food-web structure. We computed one
structural equation model to assess the relationships among the temporal trends
and one for the relationships over the spatial gradient. As spatial variation of
food-web structure are partly driven by environmental gradients, so we
controlled relationships among variables by adding two PCA axis
synthesizing respectively (1) stream gradient from upstream to downstream and
(2) enrichment (Fig. \@ref(fig:matmet)). Enrichment was characterized by the
average Biological Oxygen Demand [BOD, @mallin_factors_2006]. We further
added a random effect of the hydrographic basin on the intercept.

>>>>

---
title: "Community biomass and species richness temporal trends reveal contrasting assembly and disassembly processes in natural food-webs"
author: Alain Danet (a), Loubna El Madouri (a, c), Maud Mouchet (a), Elisa Thébault (d) and Colin Fontaine (a)
date: Research report 
output:
  #bookdown::word_document2:
  #  fontsize: 12pt
  bookdown::pdf_document2:
    fig_caption: true 
    keep_tex: true
    toc: false 
    fontsize: 12pt
header-includes:
   - \usepackage{natbib}
   - \usepackage{float}
   - \usepackage{multirow}
   - \usepackage{booktabs}
   - \usepackage{makecell}
   - \usepackage{setspace}
   - \doublespacing
   - \usepackage{lineno}
   - \linenumbers
geometry: margin=2cm
number_sections: true
bibliography: references.bib
csl: pnas.csl
---

@@

- Corresponding author: Alain Danet (a.h.danet@sheffield.ac.uk)

- Author affiliations:

(a) Centre d’Ecologie et des Sciences de la Conservation, UMR 7204 MNHN7
CNRS-Sorbonne Université, Muséum national d’Histoire naturelle de Paris, 43 rue
8 Buffon, 75005 Paris, France

(b) Department of Biosciences, Alfred Denny Building, University of Sheffield,
Western Bank, Sheffield S10 2TN, United Kingdom

(c) ?

(d) Sorbonne Université, CNRS, IRD, INRAE, Université Paris Est Créteil, 15
Institute of Ecology and Environmental Sciences of Paris (iEES-Paris), Paris,
France

- Classification: Biological Sciences (major) - Ecology (minor)

- Keywords: food-web, biodiversity changes, assembly rules, space-for-time

# Abstract (242 / 250 words max)

Understanding how biodiversity changes is a central question in ecology, but a
challenging one because of the multidimensionality of biodiversity changes.
Although the structure of ecological networks is a crucial determinant of
community dynamics, this dimension of biodiversity changes have been greatly
overlooked. One main reason is the scarce availability of the temporal series of
ecological networks. Our knowledge of the relationships among structural
properties of ecological networks comes mainly from spatial comparison, assuming
a space-for-time equivalence, which have rarely tested. We questioned the
relationships between the temporal changes of food-web structure and the ones of
community biomass and species richness, and assessed if they differed from the
ones estimated in space. We inferred the structure of using food diet and body
size data from fish monitoring in France. We estimated the temporal trends of
food-web structure in 400 sites using a hierarchical Bayesian model and then
assessed the relationships among them using structural equation models. We found
that temporal trends of connectance and average trophic level are positively
related to ones of community biomass but surprisingly poorly related to the
temporal trends of species richness, which resonates with the energetic paradigm
of food-webs. We further found general support for the space-for-time hypothesis
for community biomass, but not for species richness. Our study
sheds a new light on the dynamics of food-web assemblage in natural settings and
how the structure of ecological networks changes according to the changes of
other biodiversity facets.

# Significance (115 / 120 words max)

While the multidimensional nature of biodiversity changes has been clearly
emphasized, little is known on how the structure of food-web changes according to
changes in species richness and community biomass. By inferring food-web
structure based on food-diet data and body size ratios, we analyzed how food-web
structure in time according to concomitant changes in species richness and
community biomass. We identify that community biomass is the primary driver of
temporal changes in food-web structure, but that species richness has surprisingly
low effect. We further show that space-for-time leads to contrasting results for
species richness. Our results shed a new light on the assembly and disassembly
of food-web according to changes in biomass and in species richness.

@@

Understanding how biodiversity changes is a central question in ecology,
although challenging because biodiversity changes are multidimensional
[@mcgill_fifteen_2015; @blowes_geography_2019; @hillebrand_biodiversity_2018],
such as important changes in species composition can happen despite no changes
in species richness in average [@dornelas_assemblage_2014;
@blowes_geography_2019]. Yet species interactions are largely absent from
biodiversity changes. Although species interactions, described as ecological
networks, drive for a large part community composition, species coexistence, and
ecosystem functioning [@chesson_mechanisms_2000; @thebault_food-web_2003;
@oconnor_general_2017]. The structure of ecological networks can trigger or
impede perturbation propagations, i.e. secondary extinctions [@paine_note_1969;
@dunne_network_2002; @srinivasan_response_2007;
@stouffer_compartmentalization_2011]. Higher connectance, which measures the
degree of generalism in a network, is predicted to increase robustness to
secondary extinctions [@dunne_network_2002; @sole_complexity_2001] and studies
found that the presence of higher trophic levels is associated to a higher temporal stability of
community biomass in theoretical models and empirical data
[@shanafelt_stability_2018; @danet_species_2021]. Then, accounting for how the
structure of ecological networks changes over time according to other
biodiversity facets is crucial to understand biodiversity changes.

The availability of ecological network timeseries is scarce
[@kortsch_disentangling_2021], such as predictions about how the structure of
ecological networks change according to changes in species richness come from
spatial comparison of networks. Such studies showed that connectance is
negatively related to species richness in food-webs [reviewed in
@dunne_network_2006; @danet_species_2021], which reflects constraints on the
number of realizable species interactions or an increasing proportion of
specialists in more rich food-webs. Connectance has been shown to scale
positively or negatively with community biomass in theoretical food-webs
[@thebault_food-web_2003; @poisot_trophic_2013] while the average trophic level
was previously shown to scale negatively with species richness in empirical
food-webs [@danet_species_2021].

However, predicting how ecological networks will change in time according to
other biodiversity facets based on spatial comparison of food-webs 
is valid if the assumption of space-for-time substitution is as
well. Space-for-time substitution supposes a homology between variations in
space and in time, i.e. that spatial sequences correspond to temporal sequences.
This method raised criticisms [@damgaard_critique_2019], with studies
showing analogous patterns in space and time [@blois_space_2013] or not
[@johnson_testing_2008; @adler_contrasting_2007]. Spatial turnover of species
composition, dispersal limitation, disturbance and stochastic processes have
been hypothesized to be responsible for the non-compliance of space-for-time
substitution [@adler_contrasting_2007; @blois_space_2013].

@@

Using a Bayesian hierarchical model to estimate conjointly the temporal trends of
food-web structure at all sites, we found weak evidence of a global decrease in
community biomass (CI80%:`r gbm` by decade; median [Credible Interval], Fig. S1) but of a
global increase in species richness (CI80%: `r gr` by decade) across the `r nb_st` sites. 
We did not find evidence of a global trends in connectance and average trophic
level (CI80%: `r gct` and `r gt` respectively). The hierarchical model captured
well the variance associated to the spatial structure of the data with
conditional $R^2$ going from  `r r2ic$log_bm_std` (mean [CI95%]) for community
biomass to `r r2ic$log_rich_std` for species richness (Table S1).

@@

Using structural equation models to disentangle the direct and indirect
relationships among food-web structure descriptors, we found a strong positive
relationship between the temporal trends of community biomass and of
species richness ($r_\delta =$ `r rdtps$bm$r`, Fig. \@ref(fig:sem) a and b,
$r_\delta$: standardized slope coefficient, [95% CI]), meaning that both
temporal trends had the same direction in average. At the sites where species
richness decreased over time, community biomass also decreased over time. On the
other hand, at the sites where species richness increased over time, community
biomass also increased over time.

@@

Our results suggest that the disturbances that lead to decline or increase of
community biomass over time are concomitant to the decline and increase
respectively of trophic species at high trophic level. We found that both
connectance and average trophic level had a positive relationship with community
biomass over time, meaning that connectance and average trophic level increase
over time when community biomass increases over time as well, and likewise that
connectance and average trophic level decrease over time when community biomass
decreases over time. We also found that the temporal trends of connectance and
average trophic level were strongly correlated (Fig. S3). Likewise, we found
that the temporal trends of community biomass were positively associated to the
temporal trends of both number and proportion of piscivorous species
present (Fig. S3). This result is in accordance with the
energetic paradigm of food-webs which proposes that food-webs are primarily
constrained by fluxes of matter and energy [@lindeman_trophic-dynamic_1942],
which suggest that productivity constraints food chain length by energy
limitation [@jenkins_productivity_1992; @takimoto_ecosystem_2008]. Higher
trophic levels have been shown to be more sensible to ecosystem size
[@takimoto_ecosystem_2008; @post_ecosystem_2000], to resource availability and
disturbance in small systems [@post_long_2002] and particularly to disturbance
in hydrologic regimes in streams [@sabo_role_2010; @mchugh_dual_2010;
@jake_vander_zanden_global_2007]. In the same line, the metabolic theory of
ecology also predicts that top predators are highly constrained by energy
limitation [@arim_relationship_2007]. However, this interpretation holds if
the temporal trends of community biomass are related to the quantity of resource
available. Another explanation can be that top trophic level species are often
the ones that are targeted, harvested and those that are artificially introduced
for recreational fisheries [@reid_emerging_2019]. Further investigations on the
drivers of community biomass decline and increase over time would be needed to
further conclude.

In contrast, we found that the temporal trends of connectance was negatively
related to the temporal trends of species richness, in accordance with the
strong negative relationships that we found in the present study and from
previous results over spatial gradient [@dunne_network_2006;
@danet_species_2021]. Furthermore, we found that temporal trends of the
proportion of piscivorous species was negatively associated with the temporal
trends of species richness (Fig. S3), meaning that communities that increase in
richness over time have an increase in the proportion of non-piscivorous species over time
and vice versa. This result is in accordance with a meta-analysis showing the
number of species by trophic level decrease toward higher trophic level
[@turney_pyramids_2016]. It suggests that the temporal increase of species
richness found in our dataset is associated with a decrease of connectance
because species richness increases are biased toward non-piscivorous species
which have no piscivory links. We also found that the changes of average trophic
level over time were not related to the changes of species richness over time.
All together, an explanation can be that species richness changes over time,
without changes in community biomass, are more linked to non piscivorous species
having lower trophic level and relatively low biomass, which results in negative
relationships with temporal trends in connectance and no relationship with
average trophic level.

Our findings further suggest that species richness and community biomass have
antagonistic effects on connectance, respectively negative and positive. As we
found a strong positive relationship between the temporal trends of species
richness and community biomass, it results that species richness temporal trends
had an indirect positive effect on the temporal trends of both connectance and
average trophic level via community biomass. Then,  The temporal trends of
species richness had then a total effect on the temporal trends of connectance
and average trophic level which was twice lower than the one of the temporal
trends of community biomass. It suggests that the temporal assembly and
disassembly of food-webs are more sensitive to variations in community biomass
than in species richness, in other words that food-web structure is more robust
to changes in species richness than in biomass over time. Our study then
suggests that the changes of food-web structure over time related to changes in
species richness could be actually mediated by changes in community biomass.

As studies reported community biomass or species richness temporal trends
[@hallmann_more_2017; @ceballos_biological_2017; @dornelas_assemblage_2014;
@blowes_geography_2019], we document that temporal trends of species richness
and community biomass might have different consequences on the structure of the
community. In particular, the decrease of the average trophic level associated
to the loss of community biomass could have cascading effect on ecosystem
functioning as the highest predators are predicted to increase productivity and
stabilize communities [@wang_biodiversity_2018; @shanafelt_stability_2018;
@danet_species_2021]. In contrast, the loss of species richness might be linked
to loss of redundancy in the food-web, highlighting that those communities may
become more at risk regarding future perturbations [@yachi_biodiversity_1999].
Interestingly, the hypothesis that temporal trends in community biomass is
related to species of higher trophic level in our system has a strong reasoning
with the effect of human fisheries in marine systems which affect primarily fish
species at top trophic levels [@sprules_surfing_2016].

We found evidence that space-for-time substitution is valid for community
biomass but not for species richness. More specifically, we found a positive
relationship between community biomass and species richness both in space and
time. We also found that connectance was positively linked to community biomass
both over space and time, it resonates with theoretical studies showing that
higher food-web connectance can result in higher community biomass
[@thebault_food-web_2003], or lower [@poisot_trophic_2013] over space,
highlighting that the relation between connectance and community biomass may be
dependent on how food-web structure change with community biomass over space and
time. Likewise, we found that the average trophic level had a positive
relationship with community biomass over space and time, in accordance with a
theoretical study predicting an increase of community biomass
linked to the presence of higher predators [@wang_biodiversity_2018].
However, while we found a total negative effect between connectance and species
richness in space, in accordance with previous findings over space in
empirical food-webs [@dunne_network_2006; @danet_species_2021], we found a null
total effect in time. The result is even more contrasting for average trophic
level, where we found a total positive effect of species richness in time but a
negative one in space. This result is also in line with other studies which
found a mismatch between spatial and temporal patterns [@adler_contrasting_2007;
@johnson_testing_2008]. This discrepancy in space and time might be due to the
strong assumption that variations in space reflect differences in assembly
stages of ecological communities and then mirror their temporal
assembly/disassembly trajectories. Temporal variation of species richness might
relate to transitory dynamics while spatial gradients reflect aggregated
dynamics over a longer time scale including a far number of processes including
high species turnover along spatial gradients. Our results then suggest that food-web
structure comparison along space gradient might be a reasonable surrogate for
time sequence, it is probably not the case for species richness spatial
comparison.

@@

Our study suggests that species richness and community biomass temporal trends
in natural fish communities relate on different ecological processes. Biomass
changes in time are related to variations of food-web structure at the
higher trophic levels while changes in species richness are weakly related to
changes in food-web structure, in accordance with the energetic
paradigm of food-webs. Our results further highlight that the effects of species
richness changes on food-web structure are strongly mediated by antagonistic
effects of community biomass. We also find evidences which suggest that
space-for-time substitution may be relevant to predict the effects of community
biomass losses or gains on food-web structure but also that space-for-time
substitution might not work for species richness. Our study calls for further
investigations to support the generality of assembly rules in food-webs,
according to changes in community biomass and species richness. They are crucial
to better understand how biodiversity changes will affect community dynamics
and ecosystem functioning.

@@

We selected the sites that have been monitored at least 10 years on the
1995-2018 period, with the same sampling protocol over time. We selected
one sampling a year by site. In each site, we ensured low temporal heterogeneity
in sampling period by selecting sampling that were performed in the same period,
leading to a mean standard deviation of sampled month in each site of three
weeks [@danet_species_2021]. Those steps leads to the selection of 
`r nb_st` sites (Fig. \@ref(fig:matmet), Table S2).

@@

We first inferred the body mass of each fish individual considering an allometric
law, $B_i = 0.01 \times l_i^{3.03}$, $B_i$ being the biomass expressed in gram and $l_i$ the
body length of the individual $i$ expressed in millimeters [@froese_cube_2006]. Then the
biomass of each species and community biomass were obtained by summing the body
mass of individuals. Finally, the biomass was reported to the sampled surface
(in $m^2$) to account for sampling effort.

@@

We determined the feeding interactions of trophic species according to their
species identity and their body size. The diet database described two or three
life stages by fish species, each one being delimited by body size. The
ontogenic fish diet database was filled according to fishbase
[@froese_fishbase_2021] and literature [@bonnaffe_comparison_2021;
@danet_species_2021].

@@

At each site, yearly food-webs were described by their community biomass and
species richness. Yearly food-web structure was described by their connectance,
describing the degree of generalism in a network [@dunne_network_2006;
@delmas_analysing_2019]. It was computed as the ratio between the number of
actual links and the number of potential links ($C = \frac{L}{S(S-1)}$, with $L$
being the number of links and $S$ the number of trophic species). We also
computed the average trophic level, weighted by the biomass of the trophic
species. The trophic level of a trophic species was computed recursively as 1
added to the mean trophic level of the preys that it consumes
[@herendeen_ecological_2008].

@@

The model parameters were estimated using INLA (Integrated Nested Laplacian
Approximation) Bayesian inference with default uninformative priors
[@rue_bayesian_2017]. The prior distributions for the fixed intercept and slope were
modeled as a flat normal distribution with a mean of zero ($\mathcal{N}(0,\, 1000)$).
Prior distributions for the Gaussian error $\epsilon_{i|n, t}$ and the random
effects were modelled as a log gamma distribution with
shape and inverse scale parameters ($\mathcal{G}(s, \, \tau) = \mathcal{G}(1, \,
5.10^-5)$), then their respective standard deviation were obtained by
backtransforming the inverse scale parameter ($\sigma = 1 / \sqrt{\tau}$).
Community biomass and species richness were logged transformed.

The performance of the model was assessed visually by plotting the fitted versus
the observed values (Fig. S4). We also visually inspected the Posterior
Integral Transform (PIT) and Conditional Predictive Ordinate (CPO) to assess the
quality of the model fit. We also computed the marginal ($R^2_m$)and conditional
($R^2_c$) R-squared of the models following previous formulations
[@nakagawa_general_2013; @lahuis_explained_2014]. In order to account for the
variability associated with the priors [@gelman_r-squared_2019], we computed the
variance of the posterior distribution  of each predicted values ($Var_{fit} =
\sigma^2(\hat{Y_i})$). We computed $R^2$ values for each predicted values, then
obtaining a $R^2$ distribution [@gelman_r-squared_2019]. We reported mean
marginal and conditional $R^2$ associated the 95% credible interval computed
using the Highest Posterior Density method.

@@

We extracted the temporal trends at each site from the hierarchical model
using mean BLUP values (Best Linear Unbiaised Prediction), then multiplied
by 10 to get temporal trends by decade. For logged transformed biomass and
species richness, we derived percentage change by decade by transforming the
slopes ($\beta\prime = (e^{\beta } - 1) \times 100 \times 10$).

@@

We assessed the validity of the space-for-time substitution hypothesis by
comparing the relationships estimated among food-web structure in time to the
ones inferred in space. We used the same sites as in the temporal analysis,
but we summarised temporal series of each station by the median value over time
of each facet of food-web structure described above (Fig. \@ref(fig:matmet)).

@@

We modeled the relationships among food-web structure, that is
connectance and average trophic level, and community biomass ($dB$) and species
richness ($dR$) with Structural Equation Models. We modeled species
richness as related to community biomass and food-web structure. In
turn community biomass was related to food-web structure. We computed one
structural equation model to assess the relationships among the temporal trends
and one for the relationships over the spatial gradient. As spatial variation of
food-web structure are partly driven by environmental gradients, so we
controlled relationships among variables by adding two PCA axis
synthesizing respectively (1) stream gradient from upstream to downstream and
(2) enrichment [details in @danet_species_2021]. Enrichment was characterized by
the average Biological Oxygen Demand [BOD, @mallin_factors_2006]. We further
added a random effect of the hydrographic basin on the intercept.
